---
title: "Weather Station Data"
author: "Caroline Shepherd"
date: "2024-04-08"
output:
  html_document:
    theme: cerulean
    code_folding: hide
    highlight: monochrome
    toc: yes
    toc_depth: 5
    toc_float: yes
    collapsed: yes
    smooth_scroll: yes
    number_sections: no
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction
This code aims to visualize and synthesize data from SBBG's two weather stations. The weather stations are located within the canopy of a large Quercus agrifolia (Annie Oakly), and next to our growing oak seedlings in an open space (Oakie Dokie). 

The Oak Fuel Break experiment was designed to see if Coast Live Oak trees can provide a living fuel break, by influencing fire weather paramenters. Fire parameters include solar radiation, wind speed and direction, soil moisture, rainfall, temperature, and humidity. The following code visualizes and compares the data from two weather stations, where one is in a shaded oak area, and the other is exposed. 

```{r}

#load packages
library(tidyverse)
library(dplyr)
library(here)
library(cowplot)

```

### Initial data exploration
```{r}

# Read in the data
# Immediately changing wacky column names
annie_oakly <- read_csv(here("data", "Annie_Oakly.csv")) %>% 
  rename(id = "#",
         date_time = "Date Time, GMT-07:00",
         temp_f = "Temp, °F (LGR S/N: 21901550, SEN S/N: 21895909, LBL: Temperature)",
         rh_percent = "RH, % (LGR S/N: 21901550, SEN S/N: 21895909, LBL: RH)",
         solar_rad_wm2 = "Solar Radiation, W/m² (LGR S/N: 21901550, SEN S/N: 21903642, LBL: Solar Radiation)",
         prec_in = "Rain, in (LGR S/N: 21901550, SEN S/N: 21907472, LBL: Rain)",
         soil_m3m3 = "Water Content, m³/m³ (LGR S/N: 21901550, SEN S/N: 21907600, LBL: Water Content)",
         wind_sp_mph = "Wind Speed, mph (LGR S/N: 21901550, SEN S/N: 21907851, LBL: Wind Speed)",
         gust_sp_mph = "Gust Speed, mph (LGR S/N: 21901550, SEN S/N: 21907851, LBL: Gust Speed)",
         wind_direct = "Wind Direction, ø (LGR S/N: 21901550, SEN S/N: 21907851, LBL: Wind Direction)") 

# using lubridate to change date_time to date class
annie_oakly$date_time_local <- parse_date_time(annie_oakly$date_time, "mdy HM")

class(annie_oakly$date_time_local)

# Reading in second weather station data
oakie_dokie <- read_csv(here("data", "Oakie_Dokie_2.csv")) %>%
  rename(id = "#",
         date_time = "Date Time, GMT-08:00",
         temp_f = "Temp, °F (LGR S/N: 21901549, SEN S/N: 21895908, LBL: Temp)",
         rh_percent = "RH, % (LGR S/N: 21901549, SEN S/N: 21895908, LBL: RH)",
         solar_rad_wm2 = "Solar Radiation, W/m² (LGR S/N: 21901549, SEN S/N: 21903641, LBL: Solar Radiation)",
         prec_in = "Rain, in (LGR S/N: 21901549, SEN S/N: 21907471, LBL: Rain)",
         soil_m3m3 = "Water Content, m³/m³ (LGR S/N: 21901549, SEN S/N: 21900338, LBL: Water Content)",
         wind_sp_mph = "Wind Speed, mph (LGR S/N: 21901549, SEN S/N: 21907852, LBL: Wind Speed)",
         gust_sp_mph = "Gust Speed, mph (LGR S/N: 21901549, SEN S/N: 21907852, LBL: Gust Speed)",
         wind_direct = "Wind Direction, ø (LGR S/N: 21901549, SEN S/N: 21907852, LBL: Wind Direction)")

oakie_dokie$date_time_local <- parse_date_time(oakie_dokie$date_time, "mdy HM")

class(oakie_dokie$date_time_local)

# Initial plots for Annie Oakly
ggplot(annie_oakly, aes(x = date_time_local, y = temp_f)) +
  geom_point() +
  geom_line() +
  scale_x_datetime(breaks = "1 day") +
  theme_bw()

ggplot(annie_oakly, aes(x=date_time_local, y=prec_in))+
  geom_col()+
  scale_x_datetime(breaks="1 day")+
  theme_bw()

ggplot(annie_oakly, aes(x = date_time_local, y = soil_m3m3)) +
  geom_point() +
  geom_line() +
  scale_x_datetime(breaks = "1 day") +
  theme_bw()

# Initial plots for Oakie Dokie
ggplot(oakie_dokie, aes(x = date_time_local, y = temp_f)) +
  geom_point() +
  geom_line() +
  scale_x_datetime(breaks = "1 day") +
  theme_bw()

ggplot(oakie_dokie, aes(x=date_time_local, y=prec_in))+
  geom_col()+
  scale_x_datetime(breaks="1 day")+
  theme_bw()

ggplot(oakie_dokie, aes(x = date_time_local, y = soil_m3m3)) +
  geom_point() +
  geom_line() +
  scale_x_datetime(breaks = "1 day") +
  theme_bw()

```


### Comparisons between parameters by weather station
```{r}
#comparing soil moisture from both weather stations
soil_compare <- ggplot() +
  geom_line(data = annie_oakly, aes(x = date_time_local, y = soil_m3m3),
            color = "olivedrab4", size = .8) +
  geom_line(data = oakie_dokie, aes(x = date_time_local, y = soil_m3m3),
            color = "lightskyblue", size = .8) +
  theme_bw() +
  scale_x_datetime(breaks = "5 days") 

rain_compare <- ggplot() +
  geom_line(data = annie_oakly, aes(x = date_time_local, y = prec_in),
            color = "olivedrab4", size = .8) +
  geom_line(data = oakie_dokie, aes(x = date_time_local, y = prec_in),
            color = "lightskyblue", size = .8) +
  theme_bw() +
  scale_x_datetime(breaks = "5 days") 

plot_grid(soil_compare, rain_compare, labels = c("Soil Moisture (m3/m3)", "Precipitation (in)"), label_size = 10)


#more comparisons
wind_compare <- ggplot() +
  geom_line(data = annie_oakly, aes(x = date_time_local, y = wind_sp_mph),
            color = "blue4", size = .8) +
  geom_line(data = oakie_dokie, aes(x = date_time_local, y = wind_sp_mph),
            color = "lightskyblue4", size = .8) +
  theme_bw() +
  scale_x_datetime(breaks = "5 days") 

sun_compare <- ggplot() +
  geom_line(data = annie_oakly, aes(x = date_time_local, y = solar_rad_wm2),
            color = "red4", size = .8) +
  geom_line(data = oakie_dokie, aes(x = date_time_local, y = solar_rad_wm2),
            color = "orange2", size = .8) +
  theme_bw() +
  scale_x_datetime(breaks = "5 days") 

plot_grid(wind_compare, sun_compare, labels = c("Soil Moisture (m3/m3)", "Precipitation (in)"), label_size = 10)
  
```




